---
title: "hw-05"
author: "Will Adams"
format: html
editor: visual
embed-resources: true
---

## Question 1

```{r}
#| message: false
library(tidyverse)
exercise_data <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/visualize_data.csv")
glimpse(exercise_data)
```

I expect that people who record more exercise will have less BMI.

```{r}
cor(exercise_data$Exercise, exercise_data$BMI)

exercise_data |>
  ggplot(aes(x = Exercise, y = BMI)) +
  geom_point()
```

The result is a dinosaur drawing.

## Question 2

```{r}
#| message: false
library(causact)
glimpse(corruptDF)
```

CPI 2017 is the Corruption Perceptions Index score for different countries on a scale of 0 (highly corrupt) - 100 (very clean) for the year 2017.

HDI 2017 is the Human Development Index score for different countries based on factors including longevity, education, and income for the year 2017.

## Question 3

```{r}
corruptDF |>
  ggplot(aes(x = HDI2017, y = CPI2017)) +
  geom_point()
```

There is a positive relationship between HDI and CPI. As a country scores higher on the HDI, they appear to become more clean with a higher score on the CPI.

## Question 4

```{r}
corruptDF |>
  ggplot(aes(x = HDI2017, y = CPI2017)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "lightblue")
```

```{r}
corruptDF |>
  ggplot(aes(x = HDI2017, y = CPI2017)) +
  geom_point() +
  geom_smooth(method = "gam", se = FALSE, color = "lightgreen")
```

I prefer the gam fit, because it is more flexible in modeling the relationship between the two variables, which enables us to see how the datasets might be related in a non-linear fashion.

## Question 5

```{r}
corruptDF |>
  ggplot(aes(x = HDI2017, 
             y = CPI2017, 
             fill = region,
             color = region)) +
  geom_point() +
  facet_wrap(~region) +
  geom_smooth(method = "gam", se = FALSE)
```

The graph was quite cluttered, so I facet-wrapped by region. This shows the relationship between the CPI and HDI but segmented by region.

## Question 6

```{r}
corruptDF |>
  ggplot(aes(x = HDI2017, 
             y = CPI2017, 
             fill = region,
             color = region)) +
  geom_point() +
  facet_wrap(~region) +
  geom_smooth(method = "gam", se = FALSE) +
  scale_x_reverse()
```

## Question 7

```{r}
corruptDF |>
  ggplot(aes(x = HDI2017, 
             y = CPI2017, 
             fill = region,
             color = region)) +
  geom_point() +
  facet_wrap(~region) +
  geom_smooth(method = "gam", se = FALSE) +
  scale_x_reverse() +
  labs(x = "HDI (2017)",
       y = "CPI (2017)",
       title = "Human Development Relationship with \n Corruption Perceptions",
       subtitle = "Data points are countries",
       caption = "Source: causact")
```

## Question 8

```{r}
corruptDF |>
  ggplot(aes(x = HDI2017, 
             y = CPI2017, 
             fill = region,
             color = region)) +
  geom_point() +
  facet_wrap(~region) +
  geom_smooth(method = "gam", se = FALSE) +
  scale_x_reverse() +
  labs(x = "HDI (2017)",
       y = "CPI (2017)",
       title = "Human Development Relationship with \n Corruption Perceptions",
       subtitle = "Data points are countries",
       caption = "Source: causact")

ggsave(filename = "hw-05.png")
```
