---
title: "hw-08/09"
author: "Will Adams"
format: html
editor: visual
embed-resources: true
---

```{r}
#| message: false
library(tidyverse)
library(moderndive)
library(infer)
theme_set(theme_minimal())
```

## CH8: Polling

## Question 1

```{r}
set.seed(1108)

poll <- tibble(
  vote_gop = rbinom(n = 1000,
                    size = 1,
                    prob = .53))
```

```{r}
#First way
set.seed(1108)

bootstrap_1 <- poll |>
  rep_sample_n(size = 1000, reps = 1000, replace = TRUE) |>
  group_by(replicate) |>
  summarize(mean_gop = mean(vote_gop))

quantile(bootstrap_1$mean_gop, c(0.025, 0.975))
```

Using the first way, I got an estimated confidence interval of 0.483975 to 0.548000.

```{r}
#Second way
set.seed(1108)

bootstrap_2 <- poll |>
  specify(response = vote_gop) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "mean") |>
  get_ci(level = 0.95, type = "percentile") 

print(bootstrap_2)
```

Using the second way, I got an estimated confidence interval of 0.483975 to 0.548. These are the same because we set the seed, but if we did not there would some randomness to the simulations which would create some slight differences in the confidence intervals.

## History of Rap

```{r}
rap_poll <- read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-14/polls.csv")

rap_poll <- rap_poll |>
  filter(rank == "1")
```

## Question 2

```{r}
rap_poll |>
  ggplot(aes(x = year)) +
  geom_histogram(binwidth = 1, color = "white", center = 0) +
  labs(x = "Year of Release",
       y = "Frequency",
       title = "#1 Critic Tracks by Year of Release")
```

The year of release of the most commonly named favorite track was 1994, with 14 critics naming a track from that year.

## Question 3

```{r}
set.seed(3)

bootstrap_rap <- rap_poll |>
  rep_sample_n(size = 107, reps = 1000, replace = TRUE) |>
  group_by(replicate) |>
  summarize(mean_year = mean(year))

quantile(bootstrap_rap$mean_year, c(0.025, 0.975))
```

The lower bound of the confidence interval is 1993 and the upper bound is 1996.

## Question 4

```{r}
set.seed(3)

bootstrap_rap_2 <- rap_poll |>
  rep_sample_n(size = 25, reps = 1000, replace = TRUE) |>
  group_by(replicate) |>
  summarize(mean_year = mean(year))

quantile(bootstrap_rap_2$mean_year, c(0.025, 0.975))
```

The lower bound of the confidence interval is 1992 and the upper bound is 1998. Hence, the width of the CI is larger when we use a smaller sample because there is more room for variability in our estimates since the smaller sample might not be as representative of the larger population.

## CH9: Premier League

```{r}
#| message: false
pl_data <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/premier_league.csv")

glimpse(pl_data)
```

## Question 5

```{r}
hw_prop <- pl_data |>
  filter(result == "hw") |>
  summarize(prop_hw = n() / nrow(pl_data))

print(hw_prop)
```

The proportion of home wins was 0.4131579.

## Question 6

```{r}
set.seed(22)

sampled_proportions <- c()

for (i in 1:1000) {
  
  sampled_results <- sample(c("aw", "hw" , "d"), 
                            size = 380,
                            replace = TRUE, 
                            prob = c(1/3,1/3,1/3))
  prop <- sum(sampled_results == "hw")/380
  sampled_proportions[i] <- prop
  
}

prop_df <- data.frame(Proportions = sampled_proportions)
```

```{r}
prop_df |>
  ggplot(aes(x = Proportions)) +
  geom_histogram(binwidth = 0.01, color = "white") +
  labs(x = "Proportion of Home Wins",
       y = "Frequency",
       title = "Simulated Distribution of Home Win Proportions")
```

The histogram shows a normal simulated distribution of home wins given each result is equally as likely, centering around 0.33 (W/L/D have the same probability). The proportion of home wins in Question 5 (0.4131579) is larger than this expected value, which suggests that home wins are more frequent than if each result was equally likely.

## Question 7

Null hypothesis (H0): There is no difference in the proportion of home wins compared to other results, i.e. home / away teams win at the same rate.

Alternate hypothesis (H1): The proportion of home wins is different to other results, i.e. home / away teams win at different rates.

## Question 8

Using a two-sided hypothesis test, the p-value is the probability that the proportion of home wins is just as or more extreme than the proportion of home wins when we assume all three possible results are equally probable. Visually, we would observe two shaded portions on each tail of the null distribution of home win proportions.
