---
title: "hw-md-06-extra"
author: "Will Adams"
format: html
editor: visual
embed-resources: true
---

```{r}
#| message: false
library(tidyverse)
library(moderndive)
theme_set(theme_minimal())

data(bikes, package = "bayesrules")
glimpse(bikes)
```

## Question 0

```{r}
nrow(bikes)
```

The dataset has 500 rows. Each row represents a different day, from 2011-01-01 to 2012-12-31, although it skips some days. The highest observed ridership is 6946. The highest wind speed recorded was 34.00002 mph.

## Question 1

```{r}
get_correlation(rides ~ temp_feel,
                data = bikes) |> 
  round(2)
```

The correlation between the number of rides and what the temperature feels like is 0.58.

```{r}
get_correlation(rides ~ windspeed,
                data = bikes) |> 
  round(2)
```

The correlation between the number of rides and what the temperature feels like is -0.19.

## Question 2

```{r}
bikes <- bikes |> 
  mutate(wind_kph = (windspeed)*(1.61))

get_correlation(windspeed ~ wind_kph,
                data = bikes) |> 
  round(2)
```

The correlation between wind speed in mph and wind speed in kph is 1 since the values are simply adjusted by the 1.61 multiple we applied.
