---
title: "hw-06-1"
author: "Will Adams"
format: html
editor: visual
embed-resources: true
---

## Question 1

```{r}
#| message: false

library(tidyverse)

tv_ratings <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/tv_ratings.csv")
glimpse(tv_ratings)
```

```{r}
tv_long <- tv_ratings |>
  group_by(title) |> 
  summarise(num_seasons = n()) |> 
  ungroup() |> 
  left_join(tv_ratings, by = "title") 

tv_long <- tv_long |> 
  filter(num_seasons >= 5)

tv_long |>
  ggplot(aes(x = seasonNumber, 
             y = av_rating,
             group = title)) +
  geom_line() +
  labs(x = "Season",
       y = "Average Rating")
```

It is too messy to draw any real conclusions, except for the fact that the shows sampled have average ratings above 6, and some shows which have more than 5 seasons see a big drop in the ratings of some particular seasons.

## Question 2

```{r}
tv_long |>
  ggplot(aes(x = seasonNumber, 
             y = av_rating,
             group = title)) +
  geom_line() +
  labs(x = "Season",
       y = "Average Rating") +
  facet_wrap(~genres)
```

'Crime, Drama, and Mystery' and 'Drama, Romance' genres appear to be the two shows that lasted the longest. The show whose ratings plummeted is 'Are You Afraid of the Dark.'

## Question 3

```{r}
best_rated <- tv_ratings |>
  filter(av_rating >= 9)

best_rated |>
  ggplot(aes(x = genres)) +
  geom_bar() +
  coord_flip()
```

Coord_flip() changes the x and y axis. Drama has the most top-rated shows with 12 shows.

## Question 4

```{r}
comedies_dramas <- tv_ratings |>
  mutate(is_comedy = if_else(str_detect(genres, "Comedy"), 1, 0)) |>
  filter(is_comedy == 1 | genres == "Drama") |>
  mutate(genres = if_else(genres == "Drama","Drama","Comedy"))

glimpse(comedies_dramas)
```

```{r}
comedies_dramas |>
  ggplot(aes(x = av_rating, group = genres, color = genres)) +
  geom_density() +
  labs(x = "Average Rating",
       y = "Density",
       title = "Distribution of average ratings \n for comedies and dramas")
```

Your prediction holds somewhat true. It appears that under your classification of comedies, there is a higher proportion of comedies rated around an average rating of 8 than there are dramas, yet the peak being lower for dramas means the ratings for these shows are more spread out. Assuming the best shows have average ratings \>= 8, this suggests that the best comedies do rival the best dramas.

## Question 5

```{r}
comedies_dramas |>
  ggplot(aes(x = av_rating, group = genres, color = genres)) +
  geom_histogram() +
  labs(x = "Average Rating",
       y = "Frequency",
       title = "Distribution of average ratings \n for comedies and dramas")
```

The histogram shows us how many shows there are for each genre in this dataset - i.e. we are looking at more comedies than dramas here.

```{r}
comedies_dramas |>
  ggplot(aes(x = av_rating, group = genres, color = genres)) +
  geom_freqpoly() +
  labs(x = "Average Rating",
       y = "Frequency",
       title = "Distribution of average ratings \n for comedies and dramas")
```

The geom_freqpoly shows us the frequency of shows and their ratings as well as the distribution, allowing us to see the peaks and valleys in the distribution. I think this is the most informative because it shows both of these elements without a smooth line so we can see more data points.

## Question 6

```{r}
comedies_dramas |>
  ggplot(aes(x = av_rating, y = share)) +
  geom_bin_2d()
```

```{r}
comedies_dramas |>
  ggplot(aes(x = av_rating, y = share, fill = genres)) +
  geom_bin_2d()
```
